Homological Algebra\vspace{-1ex}

### 2019-05-22 Homological Algebra Notes/Hom Alg.tex
update

%class-specific macros
\DeclareMathOperator{\sExt}{\mathscr{E}xt}

@@

\section{May 22, 2019}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%              Homework                %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%class-specific macros
\DeclareMathOperator{\sExt}{\mathscr{E}xt}
\DeclareMathOperator{\stmod}{stmod}

@@

\section{May 22, 2019}
Some examples of triangulated categories: we have already seen that $\K(\Rmod)$ and $\D(R)$ are. But we also have
that $\stmod G,$ the category of all $G$ modules and 
\[\Hom_{\stmod G}(M,N)=\underline{\Hom}(M,N)=\Hom_G(M,N)/P\Hom_G(M,N)\]
where $P\Hom_G(M,N)$ is the collection of all morphisms that factor through a projective object $P$.

In essence, $\stmod G$ is the same as considering $G$ modules ``up to projective summand''. Recall that this makes the syzygies 
uniquely defined.

\begin{prop}
	$\stmod G$ is a tensor triangulated category.
\end{prop}
\begin{prf}
	The key idea here is that the category of $G$-modules is Frobenius, so all injectives are projective and vice versa.

	Then the triangulated structure comes from short exact sequences of $G$-modules and now our shift $\Omega$ is the syzygy operator.

	Recall that we have seen the dimension shifting identity: for $i>1$:
	\[\Ext_G^i(M,N)\cong\Ext^{i-1}(\Omega M,N)\cong\Ext^{i+1}(M,\Omega^{-1}N)\]
	the reason this doesn't actually extend to $i=1$ is that this doesn't work with $\Hom$.

	But it ends up that this actually \textit{does} work if we work with stable $\Hom$. But then we get $\Ext^1(M,N)\cong\underline{\Hom}(M,\Omega^{-1}N)$!
	so somehow this gives us a notion of a map $M\to \Omega^{-1}N$, so given the triangle $N\to P \to M$ of $G$ modules,
	\[N\to P\to M\to \Omega^{-1}N\]
	is a triangle in $\stmod G$.
\end{prf}

\subsection{Singularity category}
Let $R$ be a ring and consider $\Db(R)$, the bounded derived category. There is a subcategory $\Dbperf(R)$,
which is the bounded derived category of projective modules.

Now since the tensor of a projective module with anything is again projective, this forms a (categorical) ideal within $\Db(R)$. Taking the 
(categorical) quotient, we get $\Dsing(R)$, the \textbf{singuarity category.}

In the case that $R$ is regular (think finite global dimension), we get that any bounded complex is quasi-isomorphic to (via the Cartan-Eilenberg resolution)
to a perfect complex! So then $\Dsing(R)$ is trivial.

\begin{thm}[Rikard, Buchweitz, Orlov]
	Without assumptions on $k$ or $G$,
	\[\Dbperf(kG)\hookrightarrow \Db(kG)\to\stmod G\]
	is an exact sequence.
\end{thm}
\begin{rmk}
	Notice that this (essentially) says that semisimplicity (from the world of representation theory) is
	effectively the same as regularity (from algebraic geometry).
\end{rmk}

\subsection{Cohomological Functors}
\begin{defn}
Let $\calK$ be a triangulated category and let $\calA$ be abelian. Then $H:\calK\to \calA$ is called \textbf{cohomological}
if it takes triangles in $calK$ to long exact sequences in $\calA$.
\end{defn}
\begin{prop}
	$\Hom_\calK(X,-):\calK\to\Ab$ is cohomological.
\end{prop}
\begin{prf}
	Look at $A\xrightarrow{u}B\xrightarrow{v}C\to\Sigma A$. Now we want to show that $v_\ast u\ast=(vu)_\ast=0$
	or that $\ker v_\ast= \Im u_\ast$. Basically you can find a map to show it. Use the morphism axiom!
\end{prf}

\subsection{Yoneda Lemma}
The idea here is that objects can be determined (up to isomorphism) by the maps to or from them.

Let $\calA$ be a locally small category. Then $\mathbf{Funct}(\calA,\mathbf{Set})$ is the category of $\mathbf{Set}$-valued functors. Then there is a functor 
$\calA\to\mathbf{Funct}(\calA,\mathbf{Set})$ sending $A$ to $h_A=\Hom(A,-)$ (or $h^A=\Hom(-,A)$). Then the lemma says 
\begin{lem}[Yoneda]
	There exist bijections
	\[\Hom_\calA(A,A')\cong\mathbf{Nat}(h_{A'},h_A)\cong \mathbf{Nat}(h^A,h^{A'})\]
\end{lem}
\begin{cor}
	There is a fully faithful embedding of $\calA$ into $\mathbf{Funct}(\calA,\mathbf{Set})$.
\end{cor}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%              Homework                %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
